---
title: "Homework 4"
author: "Matt Fry"
date: "`r Sys.Date()`"
output: html_document
---

# Background/Premise

## Many of you have likely heard of the Birthday ProblemLinks to an external 
##   site. in the context of probability. In short, it says that if you have  
##   people together, and assume birthdays are equally distributed across all 
##   365 days in a year (sorry leap day people), what is the probability at 
##   least two will share a birthday? 

## It turns out, once n >= 23, this probability exceeds 0.5. 

## In this assignment, we will study the distribution of how many days match 
##    (or repeat) given a collection of  individuals. So for example, suppose 
##    6 people are in a room together, we may get the following outcomes:

# Some examples of situations

## All 6 have unique birthdays, thus 0 matches.
## 2 people share a birthday, and everyone else has a unique birthday, thus 
##    1 match.
## 3 people share a birthday, and the other 3 people have unique birthdays, 
##    thus 1 match.
## 2 people share a birthday (say March 14) and another 2 people share a 
##    birthday (say July 22), and the others have unique birthdays, thus 2 matches.
## 3 people share a birthday (say March 14) and 2 others share a different 
##    birthday (say July 22), with the sixth person different, thus 2 matches.
## 3 people share a birthday (say March 14) and the other 3 share a 
##    different birthday (say July 22), thus 2 matches.
## 2 people share a birthday (say March 14), 2 others share a different 
##    birthday (say July 22) and 2 others share yet another birthday 
##    (say Oct 31), thus 3 matches.
## 4 people share a birthday, and the other 2 people have unique birthdays, 
##    thus 1 match.
## 4 people share a birthday (say March 14), with the other 2 people sharing 
##    a different birthday (say July 22), thus 2 matches.
## 5 people share a birthday (say March 14), with the other person having 
##    a different birthday, thus 1 match.
## All 6 people share the same birthday, thus 1 match.

```{r} 
## #Modularization: getting R to generate random birthdays, using Julian Days
##  and allow for user-specified num_people; set seed for 314159
BirthdayProblem <- function(num_people=23) {
  if(!is.numeric(num_people) || num_people < 1){
    stop("num_people must be a whole, positive number")  
  } # Returns meaningful error message if num_people is not logical in the
    #   context of the problem
  else{
    set.seed(314159)
    Birthdays <- sample(1:365, size = num_people, replace = TRUE)
  }
}
```


```{r}
## #Modularization: Now, I need to add on to my original function so that
##  it provides me with the number of matching randomly generated b-days
set.seed(NULL)
BirthdayProblem <- function(num_people=23) {
  if(!is.numeric(num_people) || num_people < 1){
    stop("num_people must be a whole, positive number")  
  } # Returns meaningful error message if num_people is not logical in the
    #   context of the problem
  else{
    Birthdays <- sample(1:365, size = num_people, replace = TRUE)
    MatchingBirthdays <- num_people - length(unique(Birthdays))
  }
  print(MatchingBirthdays)
}
```


```{r}
## #Modularization: I need to run simulations now
## For 10 through 100 people in increments of 10, I will
##  run 100,000 simulations for each increment

set.seed(314159)
sapply(c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100), function(x){
  Simulation <- replicate(100000, BirthdayProblem(x))
  num_matching <- sum(Simulation > 0)
  print(paste(x, num_matching))
  summary(Simulation)
  barplot(table(Simulation))
}
)
## As you can see in the plots, as num_people increases, the number
##  of matching birthdays increases as we would expect, and the bar
##  graph becomes more normally distributed
```


# Stated Problem
## Write a function in R that randomly generates a number of birthdays and 
##    determines how many matches there are (provided above is an example). 
##    Use this function to study the distribution of the number of matches 
##    as a function of the number of people that are together.

# Coding specifics

## Your function must accept the following parameter:

## num_people corresponding to the number of people that are together, 
##    default value of 23
## Your function shall return a single value:
##    the number of days that have a match (see outline above for an example)
## Demonstrate that your function works by simulating 100,000 iterations of 
##    the process with the following details:

## Set the random number generate seed to 314159 before each run of 100,000 
##    iterations
## Run the simulation 10 times: for num_people equal to 10, 20, 30, 40, 
##    50, 60, 70, 80, 90 and 100 people.
## Provide the summary() output and a bargraph for each of the simulation runs
## Comment (as a comment or text in an Rmd) about the distribution of the 
##    number of matches



